---
title: "Simple R Functions "
author: "Kayla Ippongi"
date: "January 26, 2018"
output: pdf_document

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1.

(a) Write functions \texttt{tmpFn1} and \texttt{tmpFn2} such that if \texttt{xVec} is the vector $(x_1, x_2, ... ,x_n)$, then \texttt{tmpFn1(xVec)} returns vector $(x_1, x_2^2, ..., x_n^n)$ and \texttt{tmpFn2(xVec)} returns the vector $(x_1, \frac{x_2^2}{2}, ..., \frac{x_n^n}{n})$.

******

Here is \texttt{tmpFn1}
```{r}

tmpFn1 <- function(xVec){
  return(xVec^(1:length(xVec)))
}

## simple example
a <- c(2, 5, 3, 8, 2, 4)

b <- tmpFn1(a)
b

```

and now \texttt{tmpFn2}

```{r}
tmpFn2 <- function(xVec2){
  
  n = length(xVec2)
  
  return(xVec2^(1:n)/(1:n))
}



c <- tmpFn2(a)
c


```

######  
  
  
  
(b) Now write a fuction \texttt{tmpFn3} which takes 2 arguments $x$ and $n$ where $x$ is a single number and $n$ is a strictly positive integer.  The function should return the value of   
$$1 + \frac{x}{1} + \frac{x^2}{2} + \frac{x^3}{3} + ... + \frac{x^n}{n}$$
```{r}
tmpFn3 <- function(x,n){

  nVect = 1:n
  return(1 + sum(x^(nVect)/(nVect)))
}



c <- tmpFn3(2,5)
c


```
\newpage
### 2. Write a funciton \texttt{tmpFn(xVec)} such that if \texttt{xVec} is the vector $x = (x_1, ...,x_n)$ then \texttt{tmpFn(xVec)} returns the vector of moving averages:

$$\frac{x_1 + x_2 + x_3}{3}, \frac{x_2 + x_3 + x_4}{3}, ...,\frac{x_{n-2} + x_{n-1} + x_n}{3}$$

Try out your function.  \texttt{tmpFn}(c(1:5,6:1))
```{r}
tmpFn <- function(x){

  num = colSums(matrix(x,nrow = 3))
  denom = 3
  return(num/denom)
}

c <- tmpFn((c(1:5,6:1)))
c


```

### 3. Consider the continuous function

$$f(x) = \left\{\begin{matrix}
x^2 + 2x + 3 & if  & x < 0 \\ 
x + 3 & if & 0 \leq x < 2 \\ 
x^2 + 4x - 7 & if & 2 \leq  x 
\end{matrix}\right.$$

Write a function \texttt{tmpFn} which takes a single argument \texttt{xVec}.  the function should return the vector the values of the function $f(x)$ evaluated at the values in \texttt{xVec}.  
Hence plot the function $f(x)$ for $-3 < x <3$.
```{r}
x1 <- seq(from = -3, to = 0, length.out = 10)
x2 <- seq(from = 0, to = 2, length.out = 10)
x3 <- seq(from = 2, to = 3, length.out = 10)

y1 <- x1^2 + 2*x1 + 3
y2 <- x2 + 3
y3 <- x3^2 + 4*x3 -7

tmpFn <- function(x){
#function 1
plot(x1,y1,type = 'l',xlim=c(-3, 3), ylim=c(0, 15))
par(new = TRUE)
  
#function 2
plot(x2,y2,type = 'l', axes = FALSE ,xlim=c(-3, 3), ylim=c(0, 15))
par(new = TRUE)
  
#function 3
plot(x3,y3,type = 'l', axes = FALSE ,xlim=c(-3, 3), ylim=c(0, 15))

}
tmpFn(x)

```

### 4. Write a function which takes a single argument which is a matrix.  The function should return a matrix which is the same as the function argument but every odd number is doubled.  
Hence the result of using the function on the matrix   

$$\begin{bmatrix}
1 & 1 & 3\\ 
5 & 2 & 6\\ 
-2 & -1 & -3
\end{bmatrix}$$

should be:

$$\begin{bmatrix}
2 & 2 & 6\\ 
10 & 2 & 6\\ 
-2 & -2 & -6
\end{bmatrix}$$
```{r}
A = matrix(c(1,5,2,1,2,-1,3,6,-3),nrow = 3,ncol=3)
A
 A <- ifelse(A %% 2 !=0,A*2,A)
 A
```

### 5. Write a function which takes 2 arguements $n$ and $k$ which are positive integers.  It should return the $\mathit{n} x \mathit{n}$ matrix:


$$\begin{bmatrix}
k & 1 & 0 & 0 & \cdots  & 0 & 0\\ 
1 & k & 1 & 0 & \cdots  & 0 & 0\\
0 & 1 & k & 1 & \cdots  & 0 & 0\\
0 & 0 & 1 & k & \cdots  & 0 & 0\\
\cdot & \cdot & \cdot & \cdot & \cdot & \cdot & \cdot\\ 
0 & 0 & 0 & 0 & \cdots  & k & 1\\
0 & 0 & 0 & 0 & \cdots  & 1 & k
\end{bmatrix}$$
```{r}
tmpFn <- function(n,k){
matE <- matrix(rep(0,n*n), nrow = n, byrow = TRUE)  
matE[abs(row(matE)-col(matE))==1] <- 1
matE
diag(matE) <- k
matE
}
tmpFn(5,2)
```

### 6. Suppose an angle $\alpha$ is given as a positive real number of degrees.  
If $0 \leq \alpha < 90$ then it is quadrant 1.  If $90 \leq \alpha < 180$ then it is quadrant 2.  
if $180 \leq \alpha < 270$ then it is quadrant3.  if $270 \leq \alpha < 360$ then it is quadrant 4.  
if $360 \leq \alpha < 450$ then it is quadrant 1.  
And so on ...   
```{r}
quadrant <- function(alpha){
if(alpha < 90){
  print('quadrant 1')
}
if(alpha > 90 && alpha <= 180){
  print('quadrant 2')
}
if(alpha > 180 && alpha <= 270){
  print('quadrant 3')
}
if(alpha > 270 && alpha <= 360){
  print('quadrant 4')
}
}
quadrant(87)
```
Write a function \texttt{quadrant(alpha)} which returns the quadrant of the angle $\alpha$.


### 7.

(a) Zeller's congruence is the formula:

$$ f = ([2.6m - 0.2] + k + y + [y/4] + [c/4] - 2c)mod 7 $$
where $[x]$ denotes the integer part of $x$; for example $[7.5] = 7$.  

Zeller's congruence returns the day of the week $f$ given:

$k$ = the day of the month  
$y$ = the year in the century  
$c$ = the first 2 digits of the year (the century number)  
$m$ = the month number (where January is month 11 of the preceding year, February is month 12 of the preceding year, March is month 1, etc.)   
For example, the date 21/07/1`963 has $m = 5, k = 21, c = 19, y = 63$;  
the date 21/2/63 has $m=12, k=21, c=19, and y=62$.  
Write a function \texttt{weekday(day,month,year)} which returns the day of the week when given the numerical inputs of the day, month and year.  
Note that the value of 1 for $f$ denotes Sunday, 2 denotes Monday, etc.

```{r}

weekday <- function(day,month,year){
y <- year%%100
c <- substr(year,1, nchar(year)-2)
y <- as.numeric(y)

c <- as.numeric(c)

result <- ((2.6*month-0.2) + day + year + (y/4) + (c/4)-2*c )%%7

if(result == 1){
  print('Sunday')
}
if(result == 2){
  print('Monday')
}
if(result == 3){
  print('Tuesday')
}
if(result == 4){
  print('Wednesday')
}
if(result == 5){
  print('Thursday')
}
if(result == 6){
  print('Friday')
}
if(result == 7){
  print('Saturday')
}
}
weekday(1,2,2018)
```

(b) Does your function work if the input parameters \text{day, month, and year} are vectors with the same length and valid entries?

No it doesn't seem to work: i tested it with these vectors as parameters and got warnings because
if statements were implemented to work with 1 number not a vector of numbers. 
```{r}

x <- c(1:5)
y <- c(2,2,2,2,2)
z <- c(2018,2018,2018,2018,2018)
weekday(x,y,z)
```
  
HOWEVER, To overide these warnings i could simply put my if statments in a for loop and iterate 
through each number in the result vector

### 8.
(a)
Write a function testLoop which takes the single argument n and returns the first n  1 values of the
sequence 
```{r}
testLoop <- function(n){
  v <- c(1,2)
  for (i in n){
    if (i == 0){
      #v <- seq(v, 1)
      print(v)
      next
    }
    if(i == 1){
      #v <-seq(v,2)
      print(v)
      next
    }
    else{
      x_j = v[[i-1]] + 2/(v[[i-1]])
      v <-c(v,x_j)
    }
  }
  return(v)
}
```

(b)
Now write a function testLoop2 which takes a single argument yVec which is a vector. The function should return
```{r}
testLoop2 <- function(yVec){
  n <- length(yVec)
  result <- sum(exp(1:n))
  print(result)
  return(result)
}
```

##9 
(a)
Write a function quadmap( start, rho, niter ) 
```{r}
quadmap <- function(start,rho,niter){
  result <-c(rho*start*(1-start))
  for (i in 1:niter){
    if (i==1){
      next
    }
    else{
      #print(i)
      #print(result)
      result <-c(result,rho*result[i-1]*(1-result[i-1]))
    }
  }
  return(result)
}

tmp <- quadmap(start=0.95, rho=2.99, niter=500)
plot(tmp, type="l")
plot(tmp[300:500], type="l")
```
(b)
```{r}
fun1 <- function(start,rho){
  check = start
  count = 0
  distance = abs(start-rho)
  print(distance)
  for (i in 1:300){
    distance = distance - 0.02
    print(distance)
    count = count + 1
    print(count)
    if (distance <= 0.02){
      #print(check)
      print(count)
      return(count)
    }
  }
}
```

### 10
(a)
```{r}
xVect <- seq(from = 2, to = 56, by = 3)
tempFn<- function(xVect){
  m  = mean(xVect)
  n = length(xVect)
  r1 <- (sum(((2:n)-m) * (xVect[1])-m))/(sum((1:n)-m)^2)
  r2 <- (sum(((3:n)-m) * (xVect[1])-m))/(sum((2:n)-m)^2)
  print(r1,r2)
}
```

(b)
```{r}
xVect <- seq(from = 2, to = 56, by = 3)
tempFn<- function(xVect,k){
  m  = mean(xVect)
  n = length(xVect)
  r1 <- (sum(((2:n)-m) * (xVect[1])-m))/(sum((1:n)-m)^2)
  result <- sapply(xVect,(sum(((2:n)-m) * (xVect[1])-m))/(sum((1:n)-m)^2))
  
  print(result)
}
```
